A double-sampling extension of the German National Forest Inventory for design-based small area estimation on forest district levels

### 2018-05-10 article MDPI/hill et al 2018.tex
update 10_05_2018
update 10_05_2018

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{The German National Forest Inventory consists of a systematic grid of permanent sample plots and provides a reliable evidence-based assessment of the state and the development of Germany's forests on national and federal state level in a 10 year interval. However, the data have yet been scarcely used for estimation on smaller management levels such as forest districts due to insufficient sample sizes within the area of interests and the implied large estimation errors. In this study, we present a double-sampling extension to the existing German National Forest Inventory (NFI) that allows for the application of recently developed design-based small area regression estimators. We illustrate the implementation of the estimation procedure and evaluate its potential by the example of timber volume estimation on two small scale management levels (45 and 405 forest district units respectively) in the federal German state of Rhineland-Palatinate. An airborne laserscanning (ALS) derived canopy height model and a tree species classification map based on satellite data were used as auxiliary data in an ordinary least square regression model to produce the timber volume predictions. The results support that the suggested double-sampling procedure can substantially increase estimation precision on both management levels: the two-phase estimators were able to reduce the variance of the SRS estimator by 43\% and 25\% on average for the two management levels respectively.}

@@

The aim of this study was to investigate whether the German NFI data can provide acceptable estimation precision on two forest district levels using the latest small area estimation procedures. We therefore conducted a study in the German federal state Rhineland-Palatinate where we extended the German NFI to a double-sampling design and applied three types of design-based small area regression estimators in order to derive point and variance estimates of mean standing timber volume for 45 and 405 forest districts respectively. The SA-estimators we considered were the \textit{pseudo-small}, \textit{extended pseudo-synthetic} and the \textit{pseudo-synthetic} design-based small area estimator suggested by \citet{mandallaz2013a} and \citet{ mandallaz2013b}. Auxiliary data consisted of a canopy height model (CHM) obtained from a countrywide airborne laser scanning (ALS) and a tree species classification map to be used for regression within tree species strata. The estimation precisions were compared to those obtained by the standard one-phase estimator for cluster sampling under simple random sampling. The chosen double-sampling estimators were selected for several reasons: \textbf{(i)} the design-based framework relaxes dependencies on the regression model assumptions which seemed appropriate facing severe quality restrictions in the ALS data; \textbf{(ii)} the estimators can be used with \textit{non-exhaustive}, i.e. non wall-to-wall, auxiliary information; \textbf{(iii)} all estimators are explicitly formulated for cluster sampling which has not yet been the case for frequently used model-dependent estimators; and \textbf{(iv)} the asymptotically unbiased g-weight variance accounts for estimating the regression coefficients on the same sample used for estimation (\textit{internal model approach}) and is also robust under heteroscedasticity of the model residuals. The results from this study were considered to provide valuable information whether the suggested procedure might be a cost-saving alternative to a regional FDI.\par

@@

The one-phase estimator for cluster sampling (SRS) constitutes the \textit{status quo} that is currently applied under the existing one-phase sampling design of the German NFI in order to obtain point and variance estimates for the mean timber volume of a given estimation unit. In order to provide all estimators in the infinite population framework and ensure a consistent terminology with the two-phase estimators in Section \ref{sec:SAestimators}, we will introduce the SRS estimator that is applied in the BWI3 algorithms \citep{bwi3_ausw} in the form given in \citet{mandallaz2008, mandallaz2016}.\par 
In order to calculate the local density $Y_{c}(x)$ at the cluster level, a cluster is defined as consisting of $M$ sample locations (in the BWI3, we have $M=4$ ) where $M-1$ sample locations $x_2, ..., x_M$ are created close to the cluster origin $x_1$ by adding a fixed set of spatial vectors $e_2, ..., e_M$ to $x_1$. The actual number of plots per cluster, $M(x)$, is a random variable due to the uniform distribution of $x_l$ ($l=1, ..., M$) in the forest $F$ and to the forest/non-forest decision for each sample location $x_l$:

@@

\noindent $\pmb{Z}_{c}(x)$ is the vector of explanatory variables on the cluster level, which is calculated as the weighted average of the explanatory variables $\pmb{Z}(x_l)$ on the individual plot levels $x_1, ..., x_l$ (Eq.\ref{eq:Zc(x)}). The weight $w(x_l)$ is the proportion of the support-area within the forest $F$ used to derive the explanatory variables from the raw auxiliary information.

@@

\begin{subequations}\label{eq:compsG}
	\begin{align}
	\hat{\bar{\pmb{Z}}}_{c,G} &= \frac{\sum_{x \in s_{1,G}} M_{G}(x) \pmb{Z}_{c,G}(x)}{\sum_{x \in s_{1,G}} M_{G}(x)} \label{eq:meanZcG} \hspace{3ex} &\text{where} \hspace{3ex}
	\pmb{Z}_{c,G}(x) =  \frac{\sum_{l=1}^{L} I_{G}(x_l)\pmb{Z}(x_l)}{M_{G}(x)}\\		
	Y_{c,G}(x) &= \frac{\sum_{l=1}^{L} I_{G}(x_l)Y(x_l)}{M_{G}(x)}  &\text{and} \hspace{3ex}
	\hat{Y}_{c,G}(x)=\hat{\bar{\pmb{Z}}}_{c,G}^{\top}\hat{\pmb{\beta}}_{c,s_2}\\
	\bar{\hat{R}}_{2,G} &= \frac{\sum_{x \in s_{2,G}} M_{G}(x) \hat{R}_{c,G}(x)}{\sum_{x \in s_{2,G}} M_{G}(x)} \hspace{3ex} &\text{where} \hspace{3ex}
	\hat{R}_{c,G}(x) = Y_{c,G}(x)-\hat{Y}_{c,G}(x)
	\end{align}
\end{subequations}

Note that the restriction to $G$, i.e. $I_{G}(x_l)=\{0,1\}$, is made on the individual sample plot level $x_l$, and $M_{G}(x) = \sum_{l=1}^{L}I_{G}(x_l)$ thus is the number of sample plots per cluster within the small area. The asymptotically design-unbiased point estimate of \textit{PSMALL} is then defined according to Eq. \ref{eq:psmall_pest}. The first term estimates the small area population mean of $G$ by applying the globally derived regression coefficients to the small area cluster means of the explanatory variables $\hat{\bar{\pmb{Z}}}_{c,G}$. The second term then corrects for a potential bias of the regression model predictions in the small area $G$ by adding the mean of the empirical residuals $\bar{\hat{R}}_{2,G}$ in $G$. This correction is necessary because the zero mean residual property that holds in $F$ is not guaranteed to hold in small area $G$ under this construction.

@@

The \extpsynth{} estimator (Eq. \ref{eq:extpsynth}) has been proposed by \cite{mandallaz2013a} as a transformed version of the \psmall{} estimator that has the form of the \psynth{} estimator but remains asymptotically design unbiased. It has the advantage that the mean of the empirical model residuals of the OLS regression model for the entire area $F$ and the small area $G$ are by construction both zero at the same time, i.e. $\bar{\hat{R}}_{c} = \bar{\hat{R}}_{c,G} = 0$. This is realized by \textit{extending} the auxiliary vector $\pmb{Z}_{c}(x)$ by the indicator variable $I_{c,G}$ which takes the value 1 if the entire cluster lies within the small area $G$ and 0 if the entire cluster is outside $G$, i.e. $I_{c,G}(x)=\frac{M_{G}(x)}{M(x)}$. The extended auxiliary vector thus becomes $\pmb{\mathbb{Z}}_{c}^{\top}(x)= (\pmb{Z}_{c}^{\top}(x),I_{c,G}(x))$ and the new regression coefficient using $\pmb{\mathbb{Z}}_{c}(x)$ instead of $\pmb{Z}_{c}(x)$ in Eq. \ref{normequ_simple_cluster} is denoted as $\hat{\pmb{\theta}}_{s_2}$. All remaining components are calculated by plugging in $\pmb{\mathbb{Z}}_{c}(x)$ in Eq. \ref{eq:compsG}. A decomposition of $\hat{\pmb{\theta}}_{s_2}$ reveals that the residual correction term is now included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$.

@@

However, it is important to note that $\bar{\hat{R}}_{c,G} = 0$ under the extended regression model only holds if the sample plots $x_1, ..., x_l$ of a cluster are \textit{all} either inside our outside the small area, i.e. $M_G(x)\equiv M(x)$, and thus $I_{c,G}(x)=\frac{M_{G}(x)}{M(x)}$ can only take the values 1 or 0. \citet{mandallaz2016} assumed that the effects on the estimates should be negligible as the number of occasions where $M_{G}(x) < M(x)$ was considered to be small in practical implementations. It was thus a further objective of this study to investigate the actual occurrences and effects of this phenomenon by comparing the estimates of \extpsynth{} to those of \psmall{}. % where the estimates stay unaffected

@@

The estimation precision was quantified by the estimation error, which is the ratio of the standard error and the point estimate:

@@

We further calculated the 95\% confidence interval for each estimate for visualization purposes. The confidence intervals can also be used heuristically for hypothesis testing to determine whether the point estimates of the three estimators for a given small area are statistically different. The confidence intervals for the SRS estimator can be obtained as:

@@

In order to address the potential benefits of the small area estimators compared with the SRS approach, we calculated the \textit{relative efficiency} (Eq. \ref{eq:releff}) which can be interpreted as the relative sample size under SRS needed to achieve the variance under the double-sampling (DS) estimators.

\begin{equation}\label{eq:releff}
rel.eff=\frac{\hat{\var}(\hat{Y}_{SRS})}{\hat{\var}(\hat{Y}_{DS})}
\end{equation}

@@

Rhineland-Palatinate (RLP) is covered by a 2x2 km inventory grid of the German NFI. In the last inventory (BWI3) conducted in the year 2013, timber volume information was derived for 2810 clusters (8092 plots) in the field survey. The local timber volume density on the plot and cluster level for this sample was consequently calculated according to Section \ref{sec:srs_estimator}. In the framework of this survey, the plot center coordinates were re-measured with the differential global satellite navigation system (DGPS) technique. Knowledge about the exact plot positions were considered crucial to provide optimal comparability between the terrestrial observations and the information derived from the auxiliary information. A comparison of the DGPS coordinates with the so-far used target coordinates revealed that 90\% of all horizontal deviations lay in the range of 25 meters. A detailed analysis of horizontal DGPS errors in RLP by \citet{lambrecht2017} indicated that 80\% of the plots should not exceed horizontal DGPS errors of 8 meters. For 162 plots, the DGPS coordinates were replaced by their target coordinates due to missingness or implausible values. The terrestrial sample size $n_{2,G}$ within the FA units was 46 clusters on average and ranged between 11 and 64. Within the FR units, $n_{2,G}$ was considerably smaller with an average of 5 clusters and a range between 0 and 13.

\begin{figure}[H]
	\centering
	\resizebox{0.85\hsize}{!}{\includegraphics*{fig/SAunitsRLP.png}}
	\caption{\textit{Left}: Study area with delineated FA forest management units. \textit{Right}: Example for each of the three management units (from top to bottom): FA, FR and forest stand unit overlayed with the extended double-sampling cluster design. \textit{Green}: Forest stand polygon layer defining the forest area of this study.}
	\label{fig:StudyArean}
\end{figure}

@@

In order to apply the small area estimators (Section \ref{sec:SAestimators}), the existing NFI design was extended to a double-sampling design by densifying the existing systematic 2x2 km grid to a grid size of 500x500 m that constituted the large first phase $s_1$ in accordance to Section \ref{sec:inf_pop} (Figure \ref{fig:StudyArean}, \textit{right}). The existing terrestrial phase $s_2$ was consequently integrated by replacing the target coordinates of the respective $s_1$ clusters by the terrestrially measured DGPS coordinates. For our study, we restricted the sampling frame to the communal and state forest. The forest/non-forest decision for each plot was thereby made by a spatial intersection of the plot center coordinates with a polygon layer of the communal and state forest stands provided by the forest service. Using this stand layer provided the advantage to consistently apply the same forest/non-forest definition to the entire sample $s_1$ in order to decide about excluding or including a plot in the sampling frame. The terrestrial sample size $n_2$ was thus reduced to 2055 clusters (5791 plots). Table \ref{tab:fieldata} provides a short descriptive summary about the volume densities and the main attributes of the NFI plots located in the state and communal forest sampling frame. The densification led to an average sample size $n_{1,G}$ of 759  clusters (range: 246 -- 1022) in the FA units, and 88 clusters (range: 1 -- 194) in the FR units.

@@

A prerequisite for the application of the suggested two-phase small area estimators is the identification of suitable auxiliary data available over the entire study area. From 2003 to 2013, the topographic survey institution of RLP conducted an airborne laserscanning acquisition over the entire federal state during leaf-off conditions in order to derive a countrywide digital terrain model (DTM) as well as a digital surface model (DSM). For this study, the recorded ALS data was used to create a canopy height model (CHM) in raster format, providing discrete information about the canopy surface height of the forest area in a spatial resolution of 5 meters (Fig. \ref{fig:Auxvars}, \textit{top}). The CHM was calculated as the difference between the digital terrain model and the digital surface model that were derived by a Delauney interpolation of the ground and first ALS pulses respectively. A more detailed description of the procedure can be found in \citet{hill2017a}. The CHM provided the most valuable information to be used in the OLS regression model for predicting the timber volume on the plot and cluster level. However, it should be noted that the prolonged acquisition period of the ALS campaign led to the possibility of poor temporal alignment with the BWI3 survey, sometimes up to 10 years. In addition, the quality of the CHM varied substantially as ALS technology evolved over the years.  For example, the ALS acquisitions recorded in 2002 and 2003 exhibited particularly poor quality with about only 0.04 point per m$^2$, whereas more recent datasets contained more than 5 points per m$^2$. Furthermore, CHM information was not available at 16 sample locations due to sensor failures. These plots were deleted from the sampling frame and treated as missing at random. This assumption was considered to be reasonable as the respective sample locations did not exclude specific forest structures.

@@

% neu:
\begin{table}[H]
	\begin{center}
		\caption{Sample size for each phase in entire study area. $n_{\{1,2\},plots}$: number of plots. $n_{\{1,2\}}$: number of clusters. TSPEC: tree species map information.}
		\vspace{0.2cm}
		\label{tab:ssize}
		{\small %
			\begin{tabular}{l|r|r|r|r}
				\hline
				\multicolumn{1}{c|}{\textit{Sampling frame}} & \multicolumn{1}{c|}{\textbf{$n_{1,plot}$}}  & \multicolumn{1}{c|}{\textbf{$n_1$}}  & \multicolumn{1}{c|}{\textbf{$n_{2,plot}$}}  & \multicolumn{1}{c}{\textbf{$n_2$}} \\ % <-- added & and content for each column
				\hline \hline
				communal and state forest & 96'854 & 33'365 & 5791 & 2055\\
				\hspace{5mm} \footnotesize missing CHM & \footnotesize 18 & \footnotesize 10 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing TSPEC & \footnotesize 7060  & \footnotesize 3587 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{and} TSPEC & \footnotesize 3 & \footnotesize 2 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{or} TSPEC & \footnotesize 7075 & \footnotesize 3595 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hline \hline
			\end{tabular}
		} %
	\end{center}
\end{table}

@@

% Graphik:
\begin{figure}[H]
	\centering
	\resizebox{0.9\hsize}{!}{\includegraphics*{fig/Auxvars_400dpi.png}}
	\caption{Left: CHM (top) and tree species classification map (bottom) available on the federal state level. Right: Magnified illustration of the supports used to derive the explanatory variables from the auxiliary data.}
	\label{fig:Auxvars}
\end{figure}

@@

The tree species map was used to predict the main tree species at each sample plot which served as an additional categorical variable called \textit{treespecies}. This involved two consecutive processing steps. In the first step, one of the five tree species was assigned to a sample location if 100\% of the raster values within the edge-corrected support were classified as that species. Otherwise, the sample location was assigned the value 'mixed'. Likewise for the CHM variables, the support radius was 12 meters although the use of different support sizes for each explanatory variable would be in agreement with the two-phase estimators presented in Section \ref{sec:SAestimators}. When using the \textit{treespecies} variable in a regression model, the support size and the percentage threshold parameters had to be optimized in order to minimize the variance within each level which subsequently leads to improved model precision. A detailed analysis and description of the optimal parameter processing for the explanatory variables of the present data set is provided in \citet{hill2017a}. In a second step, the \textit{treespecies} variable was also passed through a calibration model in order to reduce the effects of misclassification errors on the regression model coefficients and to increase model accuracy. The calibration model consisted of a decision tree from a random forest algorithm \citep{breiman2001} that was trained to predict the actual main plot tree species (known for all terrestrial plots) based on available auxiliary variables. These variables were the predicted \textit{treespecies} variable, the mean canopy height and standard deviation of the CHM, as well as the proportion of coniferous trees estimated from the classification map and the growing region derived from a polygon map. The algorithm was grown with 2000 trees considering 3 of the predictors for each split. We thus applied this calibration model to the \textit{treespecies} variable derived at all sample locations $s_1$. Table \ref{tab:classacc} gives the classification accuracies \citep{congalton2008} of the \textit{treespecies} variable after calibration.

@@

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Estimation error}
\label{sec:esterr}

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{The German National Forest Inventory consists of a systematic grid of permanent sample plots and provides a reliable evidence-based assessment of the state and the development of Germany's forests on national and federal state level in a 10 year interval. However, the data have yet been scarcely used for estimation on smaller management levels such as forest districts due to insufficient sample sizes within the area of interests and the implied large estimation errors. In this study, we present a double-sampling extension to the existing German National Forest Inventory (NFI) that allows for the application of recently developed design-based small area regression estimators. We illustrate the implementation of the estimation procedure and evaluate its potential for future large scale operational application by the example of timber volume estimation on two small scale management levels (45 and 405 forest district units respectively) over the entire area of the federal German state of Rhineland-Palatinate. An airborne laserscanning (ALS) derived canopy height model and a tree species classification map based on satellite data were used as auxiliary data in an ordinary least square regression model to produce the timber volume predictions. The results support that the suggested double-sampling procedure can substantially increase estimation precision on both management levels: the two-phase estimators were able to reduce the variance of the SRS estimator by 43\% and 25\% on average for the two management levels respectively.}

@@

The objective of this study was to investigate whether the application of latest design-based small area estimation methods allow to use the German NFI data to produce estimates of acceptable precision on two forest district levels. The methods were tested in the German federal state Rhineland-Palatinate. Three types of model-assisted design-based small area regression estimators were used to derive point and variance estimates of mean standing timber volume for 45 and 405 forest management units on the two respective district levels. The SA-estimators we considered were the \textit{pseudo-small}, \textit{extended pseudo-synthetic} and the \textit{pseudo-synthetic} design-based small area estimator suggested by \citet{mandallaz2013a} and \citet{ mandallaz2013b}. Auxiliary data consisted of a canopy height model (CHM) obtained from a countrywide airborne laser scanning (ALS) and a tree species classification map to be used for regression within tree species strata. The estimation precisions were compared to those obtained by the standard one-phase estimator for cluster sampling under simple random sampling. The chosen double-sampling estimators were selected for several reasons: \textbf{(i)} the design-based framework relaxes dependencies on the regression model assumptions which seemed appropriate facing severe quality restrictions in the ALS data; \textbf{(ii)} the estimators can be used with \textit{non-exhaustive}, i.e. non wall-to-wall, auxiliary information; \textbf{(iii)} all estimators are explicitly formulated for cluster sampling which has not yet been the case for frequently used model-dependent estimators; and \textbf{(iv)} the asymptotically unbiased g-weight variance partially accounts for estimating the regression coefficients on the same sample used for estimation (\textit{internal model approach}) and is also robust under heteroscedasticity of the model residuals. The results from this study were considered to provide valuable information about the potential of the suggested small area estimation procedure and the incorporated auxiliary information for future operational large scale application.\par

@@

The one-phase estimator for cluster sampling (SRS) constitutes the \textit{status quo} that is currently applied under the existing one-phase sampling design of the German NFI in order to obtain point and variance estimates for the mean timber volume of a given estimation unit. In order to provide all estimators in the infinite population framework and ensure a consistent terminology with the two-phase estimators in Section \ref{sec:SAestimators}, we will introduce the SRS estimator that is applied in the BWI3 algorithms \citep{bwi3_ausw} in the form given in \citet{mandallaz2008} and \citet{mandallaz2016}.\par 
In order to calculate the local density $Y_{c}(x)$ at the cluster level, a cluster is defined as consisting of $M$ sample locations (in the BWI3, we have $M=4$ ) where $M-1$ sample locations $x_2, ..., x_M$ are created close to the cluster origin $x_1$ by adding a fixed set of spatial vectors $e_2, ..., e_M$ to $x_1$. The actual number of plots per cluster, $M(x)$, is a random variable due to the uniform distribution of $x_l$ ($l=1, ..., M$) in the forest $F$ and to the forest/non-forest decision for each sample location $x_l$:

@@

\noindent $\pmb{Z}_{c}(x)$ is the vector of explanatory variables on the cluster level, which is calculated as the weighted average of the explanatory variables $\pmb{Z}(x_l)$ on the individual plot levels $x_1, ..., x_l$ (Eq.\ref{eq:Zc(x)}). The weight $w(x_l)$ is the proportion of the extraction area (support) within the forest $F$ used to derive the explanatory variables from the raw auxiliary information.

@@

\begin{subequations}\label{eq:compsG}
	\begin{align}
	\hat{\bar{\pmb{Z}}}_{c,G} &= \frac{\sum_{x \in s_{1,G}} M_{G}(x) \pmb{Z}_{c,G}(x)}{\sum_{x \in s_{1,G}} M_{G}(x)} \label{eq:meanZcG} \hspace{3ex} &\text{where} \hspace{3ex}
	\pmb{Z}_{c,G}(x) =  \frac{\sum_{l=1}^{M} I_{G}(x_l)\pmb{Z}(x_l)}{M_{G}(x)}\\		
	Y_{c,G}(x) &= \frac{\sum_{l=1}^{M} I_{G}(x_l)Y(x_l)}{M_{G}(x)}  &\text{and} \hspace{3ex}
	\hat{Y}_{c,G}(x)=\hat{\bar{\pmb{Z}}}_{c,G}^{\top}\hat{\pmb{\beta}}_{c,s_2}\\
	\bar{\hat{R}}_{2,G} &= \frac{\sum_{x \in s_{2,G}} M_{G}(x) \hat{R}_{c,G}(x)}{\sum_{x \in s_{2,G}} M_{G}(x)} \hspace{3ex} &\text{where} \hspace{3ex}
	\hat{R}_{c,G}(x) = Y_{c,G}(x)-\hat{Y}_{c,G}(x)
	\end{align}
\end{subequations}

Note that the restriction to $G$, i.e. $I_{G}(x_l)=\{0,1\}$, is made on the individual sample plot level $x_l$, and $M_{G}(x) = \sum_{l=1}^{M}I_{G}(x_l)$ thus is the number of sample plots per cluster within the small area. The asymptotically design-unbiased point estimate of \textit{PSMALL} is then defined according to Eq. \ref{eq:psmall_pest}. The first term estimates the small area population mean of $G$ by applying the globally derived regression coefficients to the small area cluster means of the explanatory variables $\hat{\bar{\pmb{Z}}}_{c,G}$. The second term then corrects for a potential bias of the regression model predictions in the small area $G$ by adding the mean of the empirical residuals $\bar{\hat{R}}_{2,G}$ in $G$. This correction is necessary because the zero mean residual property that holds in $F$ is not guaranteed to hold in small area $G$ under this construction.

@@

The \extpsynth{} estimator (Eq. \ref{eq:extpsynth}) has been proposed by \citet{mandallaz2013a} as a transformed version of the \psmall{} estimator that has the form of the \psynth{} estimator but remains asymptotically design unbiased. It has the advantage that the mean of the empirical model residuals of the OLS regression model for the entire area $F$ and the small area $G$ are by construction both zero at the same time, i.e. $\bar{\hat{R}}_{c} = \bar{\hat{R}}_{c,G} = 0$. This is realized by \textit{extending} the auxiliary vector $\pmb{Z}_{c}(x)$ by the indicator variable $I_{c,G}$ which takes the value 1 if the entire cluster lies within the small area $G$ and 0 if the entire cluster is outside $G$, i.e. $I_{c,G}(x)=\frac{M_{G}(x)}{M(x)}$. The extended auxiliary vector thus becomes $\pmb{\mathbb{Z}}_{c}^{\top}(x)= (\pmb{Z}_{c}^{\top}(x),I_{c,G}(x))$ and the new regression coefficient using $\pmb{\mathbb{Z}}_{c}(x)$ instead of $\pmb{Z}_{c}(x)$ in Eq. \ref{normequ_simple_cluster} is denoted as $\hat{\pmb{\theta}}_{s_2}$. All remaining components are calculated by plugging in $\pmb{\mathbb{Z}}_{c}(x)$ in Eq. \ref{eq:compsG}. A decomposition of $\hat{\pmb{\theta}}_{s_2}$ reveals that the residual correction term is now included in the regression coefficient $\hat{\pmb{\theta}}_{s_2}$ \citep{mandallaz2016}.

@@

However, it is important to note that $\bar{\hat{R}}_{c,G} = 0$ under the extended regression model only holds if the sample plots $x_1, ..., x_l$ of a cluster are \textit{all} either inside our outside the small area, i.e. $M_G(x)\equiv M(x)$, and thus $I_{c,G}(x)=\frac{M_{G}(x)}{M(x)}$ can only take the values 1 or 0. \citet{mandallaz2016} assumed that the effects on the estimates should be negligible as the number of occasions where $M_{G}(x) < M(x)$ was considered to be small in practical implementations. It was thus a further objective of this study to investigate the actual number of occurrences as well as effects of this phenomenon by comparing the estimates of \extpsynth{} to those of \psmall{}. % where the estimates stay unaffected

@@

The estimation precision was quantified by the estimation error, which is the ratio of the standard error and the point estimate (here $\hat{Y}$ stands for the point estimate produced under the various estimators):

@@

We further calculated the 95\% confidence interval for each estimate. The confidence intervals were used heuristically for hypothesis testing to determine whether the point estimates of the three estimators for a given small area were statistically different. The confidence intervals for the SRS estimator can be obtained as:

@@

In order to address the potential benefits of the small area estimators compared with the SRS approach, we calculated the \textit{relative efficiency} (\textit{RE}, Eq. \ref{eq:releff}) which can be interpreted as the relative sample size under SRS needed to achieve the variance under the double-sampling (DS) estimators.

\begin{equation}\label{eq:releff}
RE=\frac{\hat{\var}(\hat{Y}_{SRS})}{\hat{\var}(\hat{Y}_{DS})}
\end{equation}

where $\hat{Y}$ stands for the point estimate produced under the respective estimator.

@@

Rhineland-Palatinate (RLP) is covered by a 2x2 km inventory grid of the German NFI. In the last inventory (BWI3) conducted in the year 2011 and 2012, timber volume information was derived for 2810 clusters (8092 plots) in the field survey. The local timber volume density on the plot and cluster level for this sample was consequently calculated according to Section \ref{sec:srs_estimator}. In the framework of this survey, the plot center coordinates were re-measured with the differential global satellite navigation system (DGPS) technique. Knowledge about the exact plot positions were considered crucial to provide optimal comparability between the terrestrial observations and the information derived from the auxiliary information. A comparison of the DGPS coordinates with the so-far used target coordinates revealed that 90\% of all horizontal deviations lay in the range of 25 meters. A detailed analysis of horizontal DGPS errors in RLP by \citet{lambrecht2017} indicated that 80\% of the plots should not exceed horizontal DGPS errors of 8 meters. For 162 plots, the DGPS coordinates were replaced by their target coordinates due to missingness or implausible values. The terrestrial sample size $n_{2,G}$ within the FA units was 46 clusters on average and ranged between 11 and 64. Within the FR units, $n_{2,G}$ was considerably smaller with an average of 5 clusters and a range between 0 and 13.

\begin{figure}[H]
	\centering
	\resizebox{0.85\hsize}{!}{\includegraphics*{fig/SAunitsRLP.png}}
	\caption{\textit{Left}: Study area with delineated FA forest management units. \textit{Right}: Example for each of the three management units (from top to bottom): FA, FR and forest stand unit overlayed with the extended double-sampling cluster design. \textit{Green}: Forest stand polygon layer defining the state and communal forest area of this study.}
	\label{fig:StudyArean}
\end{figure}

@@

In order to apply the small area estimators (Section \ref{sec:SAestimators}), the existing NFI design was extended to a double-sampling cluster design by densifying the existing systematic 2x2 km grid to a grid size of 500x500 m that constituted the large first phase $s_1$ (Figure \ref{fig:StudyArean}, \textit{right}). The existing terrestrial phase $s_2$ was integrated by replacing the target coordinates of the respective $s_1$ clusters by the terrestrially measured DGPS coordinates. The sampling frame was further restricted to the communal and state forest area. The forest/non-forest decision for each plot was thereby made by a spatial intersection of the plot center coordinates with a polygon layer of the communal and state forest stand layer provided by the forest service. Using this stand layer provided the advantage to consistently apply the same forest/non-forest definition to the entire sample $s_1$ in order to decide about excluding or including a plot in the sampling frame. The terrestrial sample size $n_2$ was thus reduced to 2055 clusters (5791 plots). Table \ref{tab:fieldata} provides a short descriptive summary about the volume densities and the main attributes of the NFI plots located in the state and communal forest sampling frame. The densification led to an average sample size $n_{1,G}$ of 759  clusters (range: 246 -- 1022) in the FA units, and 88 clusters (range: 1 -- 194) in the FR units.

@@

A prerequisite for the application of the suggested two-phase small area estimators is the identification of suitable auxiliary data available over the entire study area. From 2003 to 2013, the topographic survey institution of RLP conducted an airborne laserscanning acquisition over the entire federal state during leaf-off conditions in order to derive a countrywide digital terrain model (DTM) as well as a digital surface model (DSM). For this study, the recorded ALS data was used to create a canopy height model (CHM) in raster format, providing discrete information about the canopy surface height of the forest area in a spatial resolution of 5 meters (Fig. \ref{fig:Auxvars}, \textit{top}). The CHM was calculated as the difference between the digital terrain model and the digital surface model that were derived by a Delauney interpolation of the ground and first ALS pulses respectively. A more detailed description of the procedure can be found in \citet{hill2017a}. The CHM provided the most valuable information to be used in the OLS regression model for predicting the timber volume on the plot and cluster level. However, it should be noted that the prolonged acquisition period of the ALS campaign led to the possibility of poor temporal alignment with the BWI3 survey, sometimes up to 10 years. In addition, the quality of the CHM varied substantially as ALS technology evolved over the years.  For example, the ALS acquisitions recorded in 2002 and 2003 exhibited particularly poor quality with about only 0.04 point per m$^2$, whereas more recent datasets contained more than 5 points per m$^2$. Furthermore, CHM information was not available at 16 sample locations due to sensor failures. These plots were deleted from the sampling frame and treated as missing at random. This assumption was considered to be reasonable as the respective sample locations did not systematically exclude specific forest structures.

@@

% neu:
\begin{table}[H]
	\begin{center}
		\caption{Sample size for each phase in entire study area. $n_{\{1,2\},plot}$: number of plots. $n_{\{1,2\}}$: number of clusters. TSPEC: tree species map information.}
		\vspace{0.2cm}
		\label{tab:ssize}
		{\small %
			\begin{tabular}{l|r|r|r|r}
				\hline
				\multicolumn{1}{c|}{\textit{Sampling frame}} & \multicolumn{1}{c|}{\textbf{$n_{1,plot}$}}  & \multicolumn{1}{c|}{\textbf{$n_1$}}  & \multicolumn{1}{c|}{\textbf{$n_{2,plot}$}}  & \multicolumn{1}{c}{\textbf{$n_2$}} \\ % <-- added & and content for each column
				\hline \hline
				communal and state forest & 96'854 & 33'365 & 5791 & 2055\\
				\hspace{5mm} \footnotesize missing CHM & \footnotesize 18 & \footnotesize 10 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing TSPEC & \footnotesize 7060  & \footnotesize 3587 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{and} TSPEC & \footnotesize 3 & \footnotesize 2 & \footnotesize 0 & \footnotesize 0\\ % <--
				\hspace{5mm} \footnotesize missing CHM \textit{or} TSPEC & \footnotesize 7075 & \footnotesize 3595 & \footnotesize 414 & \footnotesize 385\\ % <--
				\hline \hline
			\end{tabular}
		} %
	\end{center}
\end{table}

@@

% Graphik:
\begin{figure}[H]
	\centering
	\resizebox{0.9\hsize}{!}{\includegraphics*{fig/Auxvars_400dpi.png}}
	\caption{Left: CHM (top) and tree species classification map (bottom) available on the federal state level. Right: Magnified illustration of the supports used to derive the explanatory variables from the auxiliary data. From top to bottom: CHM, aerial image, tree species classification.}
	\label{fig:Auxvars}
\end{figure}

@@

The tree species map was used to predict the main tree species at each sample plot which served as an additional categorical variable \textit{treespecies} in the regression model. In the first step, one of the five tree species was assigned to a sample location if 100\% of the raster values within the edge-corrected support were classified as that species. Otherwise, the sample location was assigned the value 'mixed'. Likewise for the CHM variables, the support radius was 12 meters although the use of different support sizes for each explanatory variable would be in agreement with the two-phase estimators presented in Section \ref{sec:SAestimators}. The specific setting for the support size and the percentage threshold was found to be optimal in order to yield the best possible regression model precision when incorporating the \textit{treespecies} variable as an additional predictor. In a second step, the \textit{treespecies} variable was also passed through a calibration model in order to reduce the effects of misclassification errors on the regression model coefficients and to increase model accuracy. The calibration model consisted of a decision tree from a random forest algorithm \citep{breiman2001} that was trained to predict the actual main plot tree species (known for all terrestrial plots) based on available auxiliary variables. These variables were the predicted \textit{treespecies} variable, the mean canopy height and standard deviation of the CHM, as well as the proportion of coniferous trees estimated from the classification map and the growing region derived from a polygon map. The algorithm was grown with 2000 trees considering 3 of the predictors for each split. We thus applied this calibration model to the \textit{treespecies} variable derived at all sample locations $s_1$. Table \ref{tab:classacc} gives the classification accuracies \citep{congalton2008} of the \textit{treespecies} variable after calibration. More details on the processing of the explanatory variables and identification of optimal parameter settings for their calculation are described in \citet{hill2017a}.

@@

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Estimation errors}
\label{sec:esterr}
